Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-08-15 paper/paper.tex
More tweaking

\begin{abstract}
    We discuss our experiences adapting three recent algorithms for maximum common (connected)
    subgraph problems to exploit multi-core parallelism. These algorithms are heavily irregular and
    very sensitive to search order changes, and so do not easily lend themselves to parallel search.
    Nonetheless, our results show that each algorithm can be parallelised successfully, with the
    multi-core algorithms we create being clearly better than the sequential ones. We then look in
    more detail at the results, and discuss how we should be
    measuring speedups for this kind of algorithm. Because of the difficulty in measuring an
    average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
    propose a new measure called \emph{aggregate speedup}.
\end{abstract}

@@

@@

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules or reactions, and also in computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, the analysis of programs and
malware \citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, and social network analysis
\citep{DBLP:journals/tkde/FangYZZ15}. A common variant of the problem requires a largest
\emph{connected} subgraph
\citep{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be inferred
from the maximum common subgraph.

@@

A \emph{clique} in a graph is a subgraph where every vertex is adjacent to every other. There is a
well-known reduction from the maximum common subgraph problem to the problem of finding a maximum
clique in an \emph{association graph}
\citep{o:Levi73,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined with
a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best approach for
solving the problem on labelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. A modified clique-like
algorithm can also be used to solve the maximum common connected subgraph problem, by ensuring
connectedness during search \citep{DBLP:conf/cp/McCreeshNPS16}; again, this is the best known way of
solving the problem on labelled graphs.

A major downside of the maximum clique approach is that the association graph encoding is extremely
memory-intensive, limiting its practical use to pairs of graphs with no more than a few hundred
vertices.

@@

However, although they are good at showing general trends, cumulative plots can hide interesting
details. We therefore now take a closer look at each of the three algorithms.

@@

We have parallelised three state-of-the-art maximum common (connected) subgraph algorithms with a
reasonable degree of success by using dynamic work-splitting. The parallel algorithms are not just
better in aggregate, but also preserve the desirable reproducibility properties of sequential
algorithms. A large part of our success was down to using parallelism to explicitly introduce
diversity into the search process, offsetting weak early value-ordering branching choices.

Despite having a branch and bound flavour, all three of the sequential algorithms had their own
difficulties and performance characteristics which prevented them from cleanly fitting into common
abstraction frameworks.

There is room for improvement, particularly with respect to work balance. However, improvements to
work balance must not come at the expense of the search order properties, nor at the cost of
increased overheads. We would also like to see distributed versions of these algorithms.

We introduced the idea of \emph{aggregate speedups}, to deal with measuring a speedup in the
presence of anomalies. These figures informed part of our analysis, but our results highlight the
importance of viewing results in multiple ways, and in using large families of instances with
different characteristics when evaluating parallel search algorithms.

>>>>

\begin{abstract}
    We discuss our experiences adapting three recent algorithms for maximum common (connected)
    subgraph problems to exploit multi-core parallelism. These algorithms are heavily irregular and
    very sensitive to search order changes, and so do not easily lend themselves to parallel search.
    Nonetheless, our results show that each algorithm can be parallelised successfully, with the
    threaded algorithms we create being clearly better than the sequential ones. We then look in
    more detail at the results, and discuss how we should be
    measuring speedups for this kind of algorithm. Because of the difficulty in quantifying an
    average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
    propose a new measure called \emph{aggregate speedup}.
\end{abstract}

@@

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules or reactions, and also in computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, the analysis of programs and malware
\citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, crisis management
\citep{o:DelavalladeFLL16}, and social network analysis \citep{DBLP:journals/tkde/FangYZZ15}.

@@

 A common variant of the problem requires a largest \emph{connected} subgraph
 \citep{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.  As
 illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be inferred
 from the maximum common subgraph.

@@

A \emph{clique} in a graph is a subgraph where every vertex is adjacent to every other. There is a
well-known reduction from the maximum common subgraph problem to the problem of finding a maximum
clique in an \emph{association graph}
\citep{o:Levi73,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined
with a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best
approach for solving the problem on labelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. A modified
clique-like algorithm can also be used to solve the maximum common connected subgraph problem, by
ensuring connectedness during search \citep{DBLP:conf/cp/McCreeshNPS16}; again, this is the best
known way of solving the problem on labelled graphs. However, the association graph encoding is
extremely memory-intensive, limiting its practical use to pairs of graphs with no more than a few
hundred vertices.

@@

However, although they are good at showing general trends, cumulative plots can hide interesting
details. We therefore now take a closer look at each of the three algorithms in turn.

@@

We have parallelised three state-of-the-art maximum common (connected) subgraph algorithms with a
reasonable degree of success by using dynamic work-splitting. Despite having a branch and bound
flavour, all three sequential algorithms had their own difficulties and performance characteristics
which prevented them from cleanly fitting into common abstraction frameworks. Nonetheless, our
results show that the parallel algorithms are not just better in aggregate, but also preserve the
desirable reproducibility properties of sequential algorithms. A large part of our success was down
to using parallelism to explicitly introduce diversity into the search process, offsetting weak
early value-ordering branching choices.

There is room for improvement, particularly with respect to work balance. However, improvements to
work balance must not come at the expense of the search order properties, nor at the cost of
increased overheads. We would also like to see distributed versions of these algorithms, to make use
of clusters of multi-core machines.

Aside from the parallel algorithms, we introduced the idea of \emph{aggregate speedups}, to deal
with measuring a speedup in the presence of anomalies. These figures informed part of our analysis,
but our results highlight the importance of viewing results in multiple ways, and in using large
families of instances with different characteristics when evaluating parallel search algorithms.
